The Impact of Hispanic Last Names and Identity on Labor Market Outcomes

### 2023-12-24 my_paper/Hadah-last-names-draft.tex
Finish writing a new version

\begin{abstract}
\singlespacing Do people with ethnically Hispanic names face discrimination in the labor market? In this study, I analyze the causal impact of Hispanic-sounding surnames on wages, using the unique context of inter-ethnic marriages and the distinct characteristics of White and Hispanic surnames. We compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities. Our findings show a significant wage difference favoring individuals with White surnames. Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this gap can be completely attributed to differences in education levels. However, this should not be interpreted as an absence of discrimination; rather, it potentially reflects discrimination in human capital accumulation. Additionally, I explore the impact of self-identifying as Hispanic on earnings. Men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not, but again, this gap can be largely explained by educational disparities. This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes as well.

@@

@@

I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences. Thus, I do not find a significant effect of having a Hispanic last name. Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.

% Inter-ethnic children with a White father, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents. Inter-ethnic children with a Hispanic father, and thus having a Hispanic last name, earn 20\% less than children of endogamous White parents. Endogamous children with Hispanic parents earn 24\% less than children of endogamous White parents. A big part of the gaps could be explained by education. Inter-ethnic children with a White father and identify as Hispanic, and thus having a non-Hispanic last name, earn 14\% less than children of endogamous White parents. Inter-ethnic children with a Hispanic father and identify as Hispanic, and thus having a Hispanic last name, earn 18\% less than children of endogamous White parents. Endogamous children with Hispanic parents and identify as Hispanic earn 10\% less than children of endogamous White parents.

@@

I use the CPS data set to study the effect of having a Hispanic last name on a person's labor market outcomes. I take advantage of the fact that the CPS asks parents' place of birth, ethnicity, and race. The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019. Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents. The census offers a larger sample of potential parents. Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity. I employ this information to construct "synthetic parents" using a method developed by \textcite{rubinstein2014pride}. I construct the synthetic parents by linking husbands and wives in the census data to each other. I assume that parents have children between the ages of 25 and 40. I then link these synthetic parents using the birth year of child, and place of birth of parents to the year of birth of the "children" and the palce of birth of parents in the CPS sample.

@@

The distribution of the four types of children is presented in figure \ref{fig:dist} and table \ref{tab:mat1}. The majority of the sample (95.96\%) is WW children. The second biggest group is HH, which constitutes 2.69\% of the sample. Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations. Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis. The summary statistics for the children of the four types of marriages are presented in table \ref{tab:c&p1}. Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure.

On the one hand, WW children are more educated, with approximately 14 years of education for men and women. WW children have an unemployment rate of 4\% for men and women. They have log hourly earnings of 2.51 for men and 2.32 for women. Also, their annual log earnings are 10.29 for men and 9.43 for women.

On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively. HH men and women have an unemployment rate of 7\% and 6\%. Their annual log earnings are 10.01 and 9.53 for men and women, respectively.

Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children. Also, WH and HW are comparable, unlike WW and HH children. These two groups are important to the analysis. Children of HW are going to be the treatment group. HW children have a Hispanic father with a Hispanic last names. At the same time, the WH children have White fathers with a non-Hispanic, or White, last names. Summary statistics of these two groups, table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.

WH children have an average of 13.58 and 13.79 years of education for men and women. The unemployment rate for men and women is 5\%. HW men and women obtained 13.22 and 13.42 years of schooling. Men's unemployment rate is 7\% and women's 6\%.

In panel C of table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic. \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic. There is considerable attrition in Hispanic identification of inter-ethnic children. Among WH children, 74\% of men and 78\% of women identify as Hispanic. Among HW children, 83\% of men and 82\% of women identify as Hispanic. These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \autocite{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}. Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic. These 5\% of WW children are most likely third generation--- or higher--- children of immigrants. For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.

In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6). Overall, HH children do worse than WW children. HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed. HH men earn 9\% less than WW men, and HH women make 2\% less than WW women. Therefore, these two groups, HH and WW, are different on many observable qualities.

The differentials between WH and HW are not as severe as the ones between HH and WW. HW children are still worse off than WH but are much better than HH. HW men attain 0.36 fewer years of education than WH men. HW women achieve 0.37 fewer years of schooling than WH women. HW children are 1 percentage points more likely to be unemployed. HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less. These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.

In table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH. Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists. The one difference is that those who identify as Hispanic do slightly worse.

@@

I show the distribution of the four types of couples in table \ref{tab:mat2}. White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples). Hispanic husbands and wives (HH) are the second largest group making up 2\% (119,749 couples) of all couples. White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples).

I present the summary statistics of the four types of synthetic parents in table \ref{tab:synth}. WW couples have higher education, 12.75 years for husbands and 12.47 for wives. As a household, WW couples have 25.22 years of schooling. Additionally, the log total family income for a WW household is 10.72 and have 3.77 children. Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling. As a household, HH couples have 17.13 years of education, log total family income of 10.27, and have 4.23 children.

WH husbands have 11.77 years of education, while wives have an average of 10.40 years. WH household attained a total of 22.17 years of schooling in total, a log total family income of 10.55, and 3.98 children. HW husbands and wives obtained 10.25 and 11.11 years of education, respectively. HW household attained a total of 21.36 years of schooling in total, a log total family income of 10.46, and 4.15 children.

In columns 5 and 6 of table \ref{tab:synth}, I report the differences in means between HH-WW households and HW-WH households. An HH Hispanic husband attained 4.11 fewer years of education than a WW husband. While an HH wife acquired 3.98 fewer years of schooling than a White WW wife. The total educational difference between HH and WW households equals 8.09 years of education. The educational differences between HW and WH are less severe than those between HH and WW. A Hispanic husband in intermarriage attained 1.52 fewer years of schooling than a White husband in intermarriage. A White wife in intermarriage attained 0.70 years of education, more than a Hispanic wife in intermarriage. The difference in total years of education between HW and WH households is 0.81, which is an 90\% decrease from the HH--WW gap. The total family income of an HH household is 45\% lower than a WW household, while the total family income of an HW household is 10\% lower than a WH household. These differences show that couples that intermarry are better and different than those that do not.

%In table \ref{tab:p1} columns five and six, I provide the  HH-WW and HW-WH gaps. WW couples are more educated and earn more than HH couples. Also, WH husbands are more educated and earn more than HW husbands. In comparison, WH wives are more educated than HW wives. Even though White husbands in intermarriages are more educated and earn more than husbands in HW marriages, they are more comparable to each other than the non-intermarriage couples.

@@

In this section I restrict the sample to WH and HW groups. Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$. $HW_{ist}$ is an indicator variables for the type of parents person $i$ has. $X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term. The equation for this strategy is written as follows:
\begin{equation} \label{eq:1a}
Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
\end{equation}

@@

@@

In this section, I present the results from estimating the two specifications in equations \ref{eq:1a} and \ref{eq:iden}. The results are shown in tables \ref{tab:lastnamereg} and \ref{tab:identreg}. All estimations are done on a sample of White, native-born men between the ages of 25 and 40 employed for full-time full-year (FTFY) and are wage and salary workers.

I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name. A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name. In other words, by comparing inter-ethnic children, a person with a Hispanic last name earns 5 percentage points less than someone with a White last name. However, more than half of the earnings gaps could be explained by educational differences. When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap. I also find a significant earnings gap between those that identify as Hispanic.

@@

I provide the results to the estimation of equation \ref{eq:1a} in table \ref{tab:lastnamereg}. The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE). I also restrict the sample to children of HW and WH (omitted) parents. Column one in table \ref{tab:lastnamereg} is the results without controlling for education. Overall, an inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name. This gap, however, could be entirely explained by educational differences. An inter-ethnic with a Hispanic last name earns 1 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.

@@

First, in Table \ref{tab:lastnamereg-weekearm}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on the full sample. Similar to the analysis before, I find that a person with a Hispanic last name earns 4 p.p. less that a person without a Hispanic last name. This gap could be explained by Educational differences. Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Mexican Hispanics. I find that a Mexican with a Hispanic sounding last name earns 3 p.p. less than a Mexican with a native sounding last name. The gap becomes an imprecise zero after controlling for education. Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of non-Mexican Hispanics. The gaps between non-Mexicans with Hispanic sounding last names and those with native sounding last names were all negative and statistically insignificant. I find that a non-Mexican with a Hispanic sounding last name earns 4 p.p. less than a Mexican with a native sounding last name. The gap becomes 1 p.p. after controlling for education. Finally, in Table \ref{tab:lastnamereg-weekearm-cub}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Cubans. I find that a Cuban with Hispanic sounding last name earns 3 p.p. less than a Cuban with a native sounding last name but the gap was statistically insignificant.

I provide the results to the estimation of equation \ref{eq:iden} in table \ref{tab:identreg}. The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE). I also restrict the sample to children of HW and WH parents. The omitted group is children of WH parents that do not identify as Hispanic. Column one in table \ref{tab:identreg} is the results without controlling for education. An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e. $HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.

@@

I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name. When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children. Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 6 percentage points less than someone with a White last name. When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap. Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name. I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.

While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination. Education itself can be influenced by bias, potentially resulting in divergent outcomes. Consequently, further research is essential to comprehensively understand the earnings gaps between Hispanics and Whites.

@@

\begin{appendices}
\section{Figures}\label{appendix:figs}

@@

\begin{center}
\begin{figure}[h!]
\caption{Distribution of the four types of children.}
\includegraphics[width=\textwidth]{PirChart2.png} 
\label{fig:dist}
\end{figure}
\end{center}

>>>>

\begin{abstract}
\singlespacing Do people with ethnically Hispanic names face discrimination in the labor market? In this study, I analyze the causal impact of Hispanic-sounding surnames on wages, using the unique context of inter-ethnic marriages and the distinct characteristics of White and Hispanic surnames. We compare the wages of individuals with one White and one Hispanic parent, observing that those with Hispanic surnames often encounter wage disparities. Our findings show a significant wage difference favoring individuals with White surnames. Although males born to Hispanic fathers and White mothers earn 5 percentage points less than those born to White fathers and Hispanic mothers, this gap can be completely attributed to differences in education levels. However, this should not be interpreted as an absence of discrimination; rather, it potentially reflects discrimination in human capital accumulation. Additionally, I explore the impact of self-identifying as Hispanic on earnings. Men with a Spanish-sounding last name who identify as Hispanic earn significantly less than those who do not, but again, this gap can be largely explained by educational disparities. This suggests that discrimination may extend beyond the labor market, influencing educational opportunities and outcomes too.

@@

The main identifying assumption of my empirical strategy depends on the assumption that people born to Hispanic fathers and White mothers are similar to their White fathers Hispanic mothers peer in all observable aspects and characteristics that are important in the labor market. Consequently, the only difference between the two groups is the variation in which group is more likely to have a Hispanic sounding last name. Children from mixed ethnic backgrounds may appear physically similar to those from single ethnicity backgrounds, but the influence of family dynamics and upbringing, crucial in developing skills and personal characteristics, varies with the pattern of mixed ethnic marriages. Particularly in a society where Hispanic ancestry is perceived negatively, it raises the question of what kind of White women would choose Hispanic men. Furthermore, even if such unions were formed randomly, children from White-Hispanic homes might benefit from more favorable family conditions than those from Hispanic-White homes, considering Whites generally have stronger socio-economic backgrounds than Hispanics. These factors introduce doubts regarding whether children from Hispanic-White families receive comparable familial support and influences, either genetically or environmentally, as those from White-Hispanic families.

@@

I find that while men born to Hispanic father-White mothers earn less than men born to White father-Hispanic mothers, this gap is entirely explained by educational differences. I find that inter-ethnic children with Hispanic sounding last name earn 5 percentage points less than those without a Hispanic last name. The gap becomes statistically insignificant after controlling for education. Thus, I do not find a significant effect of having a Hispanic last name. Finally, I find that men identifying as Hispanic earn significantly less than those who do not, even when I control for ancestry and education.

@@

I use the CPS data set to study the effect of having a Hispanic last name on a person's labor market outcomes. I take advantage of the fact that the CPS asks parents' place of birth, ethnicity, and race. The data spans the period between 1994--- the earliest sample to ask about a parent's place of birth--- to 2019. Moreover, since the CPS does not provide data on parents' characteristics, essential to determine the family background, I use the census to construct synthetic parents. The census offers a larger sample of potential parents. Similar to the CPS, the 1960 to 2000 censuses ask about the person's place of birth and the individual's race and ethnicity. I employ this information to construct "synthetic parents" using a method developed by \textcite{rubinstein2014pride}. I construct the synthetic parents by linking husbands and wives in the census data to each other. I assume that parents have children between the ages of 25 and 40. I then link these synthetic parents using the birth year of child, and parents' places of birth to the year of birth of the "children" and the parents' places of birth in the CPS sample.

@@

The distribution of the four types of children is presented Table \ref{tab:mat1}. The majority of the sample (96\%) is WW children. The second biggest group is HH, which constitutes 3\% of the sample. Inter-ethnic children, WH and HW, make up 1.35\% of the sample with 90,325 observations. Even though WH and HW are only 1.35\% of the sample, I have plenty of observations to carry out an analysis. The summary statistics for the children of the four types of marriages are presented in Table \ref{tab:c&p1}. Children of WW marriages (column 1) do better on every measure while children of HH parents (column 4) do worse than other children on every measure. Children of WH (column 2) and HW (column 3) marriages fall in between WW and HH children. Moreover, the summary statistics for those that also identify as Hispanic is reported in in Table \ref{tab:c&p2}.

% On the one hand, WW children are more educated, with approximately 14 years of education for men and women. WW children have an unemployment rate of 4\% for men and women. They have log hourly earnings of 2.51 for men and 2.32 for women. Also, their annual log earnings are 10.29 for men and 9.43 for women.

% On the other hand, children of HH marriages have an average of 12.9 and 13.24 years of education for men and women, respectively. HH men and women have an unemployment rate of 7\% and 6\%. Their annual log earnings are 10.01 and 9.53 for men and women, respectively.

% Inter-ethnic children, WH and HW, have lower educational attainment and income than WW children, but they are better than HH children. Also, WH and HW are comparable, unlike WW and HH children. These two groups are important to the analysis. Children of HW are going to be the treatment group. HW children have a Hispanic father with a Hispanic last names. At the same time, the WH children have White fathers with a non-Hispanic, or White, last names. Summary statistics of these two groups, Table \ref{tab:c&p1} columns (2) and (3) can show that they are different from the children of non-inter-ethnic marriages.

% WH children have an average of 13.58 and 13.79 years of education for men and women. The unemployment rate for men and women is 5\%. HW men and women obtained 13.22 and 13.42 years of schooling. Men's unemployment rate is 7\% and women's 6\%.

% In panel C of Table \ref{tab:c&p1}, I present the rates at which members of each group self-identify as Hispanic. \footnote{Hispanic self-identification is observed when an individual notes that they are Hispanic on the CPS questionnaire} The vast majority of HH children, 96\% for men and 97\% for women, identify as Hispanic. There is considerable attrition in Hispanic identification of inter-ethnic children. Among WH children, 74\% of men and 78\% of women identify as Hispanic. Among HW children, 83\% of men and 82\% of women identify as Hispanic. These results align with the findings of a substantial body of literature that noted this attrition among children of immigrants, especially those coming from inter-ethnic marriages \autocite{duncan2017complexity, duncan2018identifying, duncan2020new, antman2020ethnic}. Finally, a small portion of WW children, 4\% of men and 5\% of women, identified as Hispanic. These 5\% of WW children are most likely third generation--- or higher--- children of immigrants. For my analysis, I dropped the WW observations who identify as Hispanic when I compare with non-Hispanic Whites.

% In columns (5) and (6), I calculate the differences in means between HH-WW (column 5) and HW-WH (column 6). Overall, HH children do worse than WW children. HH has one year less of education compared to WW, and are 2 percentage points more likely to be unemployed. HH men earn 9\% less than WW men, and HH women make 2\% less than WW women. Therefore, these two groups, HH and WW, are different on many observable qualities.

% The differentials between WH and HW are not as severe as the ones between HH and WW. HW children are still worse off than WH but are much better than HH. HW men attain 0.36 fewer years of education than WH men. HW women achieve 0.37 fewer years of schooling than WH women. HW children are 1 percentage points more likely to be unemployed. HW men earn 1\% less annual earnings as WH men, and HW women earn 3\% less. These differences show that HW and WH children are different from children of endogamous/homogamous marriages and, thus, are more comparable.

% In Table \ref{tab:c&p2}, I present the summary statistics for the four types and limit the sample to those that only identify as Hispanic.\footnote{Limiting the sample to those who identify as Hispanic only affects WH, HW, and HH. Members of the WW group are not affected by this restriction.} The same pattern discussed above between HH--WW and HW--WH persists. The one difference is that those who identify as Hispanic do slightly worse.

@@

I show the distribution of the four types of couples in Table \ref{tab:mat2}. White husbands and White wives (WW) make up the majority of couples in the sample, 96\% (5,141,737 couples). Hispanic husbands and wives (HH) are the second largest group making up 2\% (119,749 couples) of all couples. White husbands and Hispanic wives (WH) couples are less 1\% (33,097 couples) of the sample and Hispanic husbands and White wives (HW) are also 1\% (37,847 couples). I present the summary statistics of the parents in Table \ref{tab:synth}.

@@

In this section I restrict the sample to WH and HW groups. Let $Y_{ist}$ be the annual log earnings of person $i$ in state $s$ at time $t$. $HW_{ist}$ is an indicator variables for the type of parents of person $i$ has. $X_{ist}$ is a vector of controls that includes age and numbers of hours worked, $\gamma_{t}$ are year fixed effects, $\lambda_{s}$ are state fixed effects, and $\phi_{ist}$ represents the error term. The equation for this strategy is written as follows:
\begin{equation} \label{eq:1a}
Y_{ist} = \beta_{1} HW_{ist} + X_{ist} \pi + \gamma_{t} + \lambda_s + \phi_{ist}
\end{equation}

@@

\subsection{Threats to Identification}

The central assumption underpinning my estimation strategy is based on the hypothesis that individuals born to a Hispanic father and White mother exhibit comparable characteristics to their peers of White-Hispanic descent, especially in areas vital to the labor market. This assumption includes similarities in educational background, skill sets, and work experiences that are significant determinants in employment opportunities, salary levels, and career advancement.

Two primary reasons underscore that this assumption. First, there is a significant selection in the marriage market. Since belonging to the Hispanic out-group comes with a negative penalty in society, one might wonder who are the White women that would be willing to give their kids a Hispanic last-name that would potentially negatively influence their futures. Second, fathers and mothers influence human capital accumulation differently \autocite{kimball2009risk,magruder2010intergenerational}. If marriage was random, then WH households might be a better environment for children that HA. Using the CPS and the US Census, I will evaluate these empirical concerns.

The previous two threats to identification could be addressed by the selection of partners among inter-ethnic couples. The average White person exhibits better observable characteristics than the Average Hispanic. Consequently, if marriages were random, the White husband in an WH marriage would have better traits than wife, and the White wife in the HW marriage would have the stronger traits. The random matching coupled with the fact that parents are not perfect substitutes in the inter-generational transmission of human capital would render the identification invalid since the parental background will affect labor market outcomes.

Marriages, however, are not random. There is large body of empirical and theoretical work that indicates that marriages exhibits strong selection, or what is referred to as assortative matching, on traits \autocite{averettBetterWorseRelationship2008, averettEconomicRealityBeauty1996, beckerTheoryMarriagePart1973, beckerTheoryMarriagePart1974, beckerTreatiseFamily1993, browningCollectiveUnitaryModels2006, chiapporiFatterAttractionAnthropometric2012}. Also, \textcite{duncanIntermarriageIntergenerationalTransmission2011} shows a similar pattern of assortative matching among Mexicans in the United States. These papers suggest that inter-ethnic WH and HW marriages will, on average, consist of partners sharing similar characteristics. Consequently, the literature would predict that both WH and HW parents will be similar in all aspects that are relevant to a child's labor market outcomes. I will go over the empirical evidence from the data to show that this is actually the case.

\section{From the Data: The Differences Between HW and WH Couples}\label{sec:hw-wh-couples-data}

In this section, I explore the empirical data to affirm the validity of my empirical strategy, as presented in Table \ref{tab:synth}, which details the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH). This table meticulously lays out the average outcomes and discrepancies for each group, highlighting the impact of inter-ethnic marriages on children's prospects. These results show that there is selection in marriage, and that the differences between children born to Hispanic fathers and White mother (HW) and children born to White fathers and Hispanic mother (WH) are less severe that those between the other two groups. Consequently, comparing WH and HW children to each other to analyze discrimination against Hispanics in the labor market.

Using the ``synthetic'' parents sample that I constructed using the Census data, I examine the family background of the different type of children. WW couples have higher education, 12.58 years for husbands and 12.36 for wives. As a household, WW couples have 24.95 years of schooling. Husbands in HH marriages have 8.64 years of education, while women have 8.49 years of schooling. As a household, HH couples have 17.13 years of education. As predicted, the inter-ethnic couples exhibit matching in which partners marry people similar to them. WH husbands have 11.82 years of education, while wives have an average of 10.71 years. WH household attained a total of 22.68 years of schooling in total. HW husbands have 10.33 years of education, while wives have an average of 11.01 years. HW household attained a total of 21.50 years of schooling in total. Both Hispanic men and women in inter-ethnic couples marry white spouses that are, on average, more educated the the average HH couple. More importantly for human capital accumulation, the mother in the HW marriages---the children that have a Spanish sounding last-name---are more educated than their WH peers. Since mothers could be more important in the a child's human capital accumulation and education, this would suggest that a child with a Hispanic last-name should complete more years of education.

The data reveals that WW couples have the highest total household education, amounting to 24.95 years, significantly surpassing the 17.69 years of HH couples. This stark contrast underlines the substantial educational divide between these groups. Inter-ethnic couples, WH and HW, display intermediate educational achievements, with WH households totaling 22.68 years of education and HW households slightly behind at 21.50 years. Notably, HW husbands are less educated (10.33 years) compared to WH husbands (11.82 years), yet HW wives surpass their WH counterparts with 11.01 years of education, suggesting a balance in educational attainment within these marriages. This factor is particularly pertinent for children with Hispanic last names who might derive greater benefits from their mother's higher education.

In terms of labor market performance, WW households boast the highest log total family income at 10.75, while HH households fall at the lower end with 10.42. Among inter-ethnic couples, WH households have a slightly higher total income of 10.65 compared to HW households at 10.60. Notably, the difference in husbands' log hourly earnings between HW and WH is marginal, indicating HW men earn 4\% less than their WH counterparts. Furthermore, HW women actually surpass WH women in earnings, with HW wives earning 1.75 and WH wives earning 1.73 in log hourly earnings, respectively. This reversal in the typical earning pattern not only speaks to the closing economic disparities between these groups but also implies potentially greater economic contributions from HW women to their families, which could be advantageous for children and to the developments of traits that are important in the labor market.

The table also sheds light on fertility trends, noting that WW couples have less children than HH couples, reflecting broader socio-economic and cultural patterns. The difference in fertility between HW and WH couples is positive but significantly lower than the difference in fertility between HH and WW couples.

The evidence presented in this section supports a robust empirical strategy, revealing significant selection in marriage and distinctive educational and income patterns among different types of families. Inter-ethnic couples exhibit comparable levels of education and earnings. This suggests that the children of HW families are likely positioned for better educational outcomes, critical in understanding discrimination in the labor market. Notably, the disparities between HW and WH families are considerably less pronounced than those between other groups, emphasizing the importance of comparing these children directly. Such comparisons shed light on the nuanced dynamics of ethnicity, education, and economic outcomes in inter-ethnic marriages. \footnote{I present in Tables \ref{tab:synthmex} and \ref{tab:syntnonmex} the summary statistics which detail the educational and economic profiles of parents from four different ethnic groupings—White White (WW), White Hispanic (WH), Hispanic White (HW), and Hispanic Hispanic (HH) on sub-samples of Hispanics of Mexican and non-Mexican ancestries. I find similar results that describe a selection into inter-ethnic marriages among the two groups.}

Despite higher levels of education and income among Hispanic White (HW) mothers compared to White Hispanic (WH) mothers, suggesting an expectation of greater educational attainment for HW children, data reveals a discrepancy. Specifically, HW children complete, on average, 0.4 fewer years of education compared to their WH peers (refer to Table \ref{tab:c&p1}). This gap may suggest potential discrimination or barriers in educational access for HW children.

@@

In this section, I present the results from estimating the two specifications presented in equations \ref{eq:1a} and \ref{eq:iden}. I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers. The results are shown in Tables \ref{tab:lastnamereg} and \ref{tab:identreg}.

I find that an inter-ethnic person with a Spanish-sounding last names earns less than an inter-ethnic who has a White last name. A person with a Spanish-sounding last name earns 5 percentage points less than a person with a White last name. In other words, by comparing inter-ethnic children, a person with a Hispanic last name earns 5 percentage points less than someone with a White last name. However, more than half of the earnings gaps could be explained by educational differences. When I control for education, the last name effect decreases to a statistically insignificant 1 percentage points earnings gap. I also find a significant earnings gap between those that identify as Hispanic.

@@

I provide the results to the estimation of equation \ref{eq:1a} in Table \ref{tab:lastnamereg}. I estimate the mean wages of White, U.S.-born, Hispanic men aged 25-40 who are employed full-time and full-year (FTFY) as waged and salaried workers. I also restrict the sample to children of HW and WH (omitted) parents. Column one in Table \ref{tab:lastnamereg} is the average crude earnings gap in log annual earnings between HW workers and their WH peers. In the next 4 columns, I introduce the results with controls for hours worked, state FE, year FE, age FE, Education FE, and parental background.

Overall, the crude gap between HW and WH workers is equal to 5 percentage points (Table \ref{tab:lastnamereg} column 1). An inter-ethnic with a Hispanic last name earns 5 percentage points less than an inter-ethnic with a White last name. Even after controlling for hours worked, and including state, year, and age FEs in the estimation, the gap stays at 5 percentage points. This gap, however, could be entirely explained by educational differences. An inter-ethnic with a Hispanic last name earns 1 percentage points less than an inter-ethnic with a White last name, but the result is statistically insignificant.

@@

First, in Table \ref{tab:lastnamereg-weekearm}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on the full sample. Similar to the analysis before, I find that a person with a Hispanic last name earns 4 p.p. less that a person without a Hispanic last name. This gap could be explained by Educational differences. Second, in Table \ref{tab:lastnamereg-weekearm-mex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Mexican Hispanics. I find that a Mexican with a Hispanic sounding last name earns 3 p.p. less than a Mexican with a native sounding last name. The gap becomes an imprecise zero after controlling for education and parental background. Third, in Table \ref{tab:lastnamereg-weekearm-nonmex}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of non-Mexican Hispanics. The gap between non-Mexicans with Hispanic sounding last names and those with native sounding last names was similarly explained by educational differences. I find that a non-Mexican with a Hispanic sounding last name earns 3 p.p. less than a Mexican with a native sounding last name. The gap becomes an imprecise zero after controlling for education and parental background. Finally, in Table \ref{tab:lastnamereg-weekearm-cub}, I present the results to the estimation of equation \ref{eq:1a} using weekly earnings as a dependent variable on a sample of Cubans. I find that a Cuban with Hispanic sounding last name earns 4 p.p. less than a Cuban with a native sounding last name but the gap was statistically insignificant.

I provide the results to the estimation of equation \ref{eq:iden} in Table \ref{tab:identreg}. The sample in this analysis includes full-time, full-year, and wage and salary men, and I control for hours worked and age and include year fixed effects (FE). I also restrict the sample to children of HW and WH parents. The omitted group is children of WH parents that do not identify as Hispanic. Column one in Table \ref{tab:identreg} is the results without controlling for education. An inter-ethnic $HW_{ist}$ that identifies as Hispanic \textit{versus} $HW_{ist}$ who do not---i.e. $HW_{ist} + HW_{ist} \cdot Hispanic_i$---earns 6 percentage points less than an inter-ethnic that does not, but the result could be entirely explained by educational differences.

@@

I compare the children of inter-ethnic marriages to study the labor market impact of having a Hispanic last name. When I compare the earnings of HW and WH children, which captures the effect of having a Hispanic last name, HW children earn  6 percentage points less than WH children. Thus, by comparing inter-ethnic children, a person with a Hispanic last name makes 5 percentage points less than someone with a White last name. When I control for education, the last name effect decreases to a statistically insignificant 2 percentage points earnings gap. Moreover, I estimate the effect of identifying as Hispanic and having a Hispanic last name. I find that an inter-ethnic child with a Hispanic last name and identifies as Hispanic earns 5 percentage points less that an inter-ethnic child with native sounding last name and identifies as Hispanic.

While the earnings gap between children of inter-ethnic parents with and without a Hispanic last name disappears when controlling for education, it does not necessarily indicate the absence of discrimination. Education itself can be influenced by bias, potentially resulting in divergent outcomes. This is especially the case when parental characteristics indicate that people with a Hispanic-sounding last-names should in theory complete more years of education---since mothers of inter-ethnic children with a Hispanic last-names have more years of education and earn more. Consequently, further research is needed to comprehensively understand the earnings gaps between Hispanics and Whites.

@@

\begin{appendices}
% \section{Figures}\label{appendix:figs}

@@

% \begin{center}
% \begin{figure}[h!]
% \caption{Distribution of the four types of children.}
% \includegraphics[width=\textwidth]{PirChart2.png} 
% \label{fig:dist}
% \end{figure}
% \end{center}

% \newpage
